Effect of Black Hole Spins on Christodoulou Memory in Binary Mergers

### 2019-03-27 notes/Paper.tex
some corrections

\stm{note everywhere I have added ``the'', that ``upto'' should be ``up to'', etc}
\begin{abstract}
	The Christodoulou memory, which is a nonlinear memory effect sourced by the oscillatory component of the gravitational wave stress tensor, produces a growing, nonoscillatory change in the gravitational wave \st{"plus" polarization} strain \stm{the ``plus'' polarization isn't a fundamental quantity, its a gauge choice}. This results in the permanent displacement of a pair of freely falling test masses after the gravitational wave has passed. While the memory contribution during the early inspiral is well described by the post-Newtonian (PN) approximation, the contribution at merger must be obtained using numerical relativity simulations. The corrections up to third PN order (i.~e.~3PN) for the gravitational wave memory from pairs of nonspinning black holes on quasicircular orbits has been computed by Favata et al. \stm{is it Favata et al, or just Favata by himself? use bibtex for references, get style file for PRD so that they appear correctly}. Here we include the contribution from the individual black-hole spins at leading order \stm{what is leading order here, if we then say the memory is 3PN?} and calculate the memory effect up to 3PN order. We also compute the memory contribution during
the merger by using the methods presented in \stm{cite Favata} to calculate the memory content using the oscillatory radiation as input, and
applying that method using publicly available numerical relativity simulations of spinning binaries.
\st{using Weyl curvature component $\psi_4$ after removing the unphysical low frequency contribution due to spectral leakage. We use the method described in [Class. Quantum Grav. 28 195015] for integration which control the amplification of the unphysical frequency resulting from spectral leakage.}\stm{This is too much detail form the abstract, belongs in the body of the paper} 
\end{abstract}

@@

Gravitational waves (GW), when they pass through a pair of freeling falling test masses, induce an oscillatory motion. This motion is incredibly small for astrophysical sources. Nonetheless, the signals generated by ten black-hole binaries and one neutron-star binary were detected by the Advanced LIGO and Advanced Virgo \stm{cite the catalog paper}. For sources whose strength changes significantly with time, such as the final merger of compact binaries like those that have already been observed, there is an additional non-oscillatory signal that results in a permanent displacement of
freely falling test masses. This permanent displacement, which is called the Christodoulou memory effect \cite{Christodoulou1991}, is sourced by the oscillatory part of the signal, and is therefore a fundamental manifestation of nonlinearity in general relativity. This memory effect accumulates over the entire history of the source. This nonlinear memory effect is distinct from linear memory, which is due to a change in the time derivative of one or more multipole moments of a source \stm{give a more specific definition}. The linear memory is important for unbound astrophysical systems and have been studied in context of supernova explosions \stm{put one or more citations after each item}, asymmetric mass loss due to neutrino emission \stm{isn't this also in the context of supernovae? if not, what source is this?}, and gamma-ray-burst jets \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}.

@@

This wave equation has several source terms. The Landau-Lifshitz term represents the contribution of GWs to the stress tensor,
and therefore serves as the source of the nonlinear memory contribution to GW waveform \cite{Isaacson1968},
\stm{You appear to change the meanings of T and t, which you shouldn't do, but please verify that my correction is what you actually meant. Also, since the contravariant form appears in the above equation, you should give that here, rather than the covariant form.}
\begin{equation}
	t^{LL}_{jk}\equiv \frac{1}{32\pi}\langle h^{TT}_{ab,j}h^{TT}_{ab,k}\rangle\approx t^{LL}_{00}n_{j}n_{k} = \frac{1}{R^{2}}\frac{dE^{gw}}{dt d\Omega}n_{j}n_{k}
\end{equation}
where $TT$ refers to the transverse-traceless gauge, $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-brackets correspond to an average over several wavelengths, \stm{what does this last one mean?} $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$. The standard procedure to solve a wave equation with a source is to use the Green's function. When this approach is applied to Eq. \ref{eq:3.1a}, we obtain a correction term to the gravitational field sourced by the Landau-Lifshitz term, given by \cite{Wiseman1991}
\begin{equation}\label{eq:1.6}
	\delta h^{TT}_{jk} = \frac{4}{R}\int_{-\infty}^{T_{R}} dt'\Bigg[ \frac{dE^{gw}}{dt'd\Omega'}\frac{n'_{j}n'_{k}}{(1-\boldsymbol{n'}\cdot\boldsymbol{N})}d\Omega'\Bigg]^{TT}
\end{equation}
where $T_{R}$ is the retarded time. Eq.~\ref{eq:1.6} is the nonlinear Landau-Lifshitz pseudotensor describing gravitational wave memory \cite{Christodoulou1991, Payne1894, Blanchet_Damour1992}. The nonlinear memory occurs in any system that radiates GWs, whereas these systems are usually considered to have a vanishing linear memory, although we can apply the formalism for linear memory to nonlinear memory if 
we consider the radiated gravitons to be the unbound masses \cite{Thorne1992}.
%The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992}

@@

%\bibliographystyle{apsrev4-1}
\bibliographystyle{prsty}
\bibliography{references}

>>>>

\stm{note everywhere I have added ``the'', that ``upto'' should be ``up to'', etc}
\begin{abstract}
	The Christodoulou memory, which is a nonlinear memory effect sourced by the oscillatory component of the gravitational wave stress tensor, produces a growing, nonoscillatory change in the gravitational wave strain. This results in the permanent displacement of a pair of freely falling test masses after the gravitational wave has passed. While the memory contribution during the early inspiral is well described by the post-Newtonian (PN) approximation, the contribution at merger must be obtained using numerical relativity simulations. The corrections up to third PN order (i.~e.~3PN) for the gravitational wave memory from pairs of nonspinning black holes on quasicircular orbits has been computed by Favata \cite{Favata2009}. Here we include the contribution from the individual black-hole spins up to 1.5PN \stm{what is leading order here, if we then say the memory is 3PN?} and calculate the memory effect up to 3PN order. We also compute the memory contribution during
the merger by using the methods presented in \cite{Favata2009} to calculate the memory content using the oscillatory radiation as input, and
applying that method using publicly available numerical relativity simulations of spinning binaries.
\end{abstract}

@@

Gravitational waves (GW), when they pass through a pair of freeling falling test masses, induce an oscillatory motion. This motion is incredibly small for astrophysical sources. Nonetheless, the signals generated by ten black-hole binaries and one neutron-star binary were detected by the Advanced LIGO and Advanced Virgo\cite{Ligocatalog2018} \stm{cite the catalog paper}. For sources whose strength changes significantly with time, such as the final merger of compact binaries like those that have already been observed, there is an additional non-oscillatory signal that results in a permanent displacement of
freely falling test masses. This permanent displacement, which is called the Christodoulou memory effect \cite{Christodoulou1991}, is sourced by the oscillatory part of the signal, and is therefore a fundamental manifestation of nonlinearity in general relativity. This memory effect accumulates over the entire history of the source. \ashok{This nonlinear memory effect is distinct from linear memory, which arises from the non-oscillatory motion of unbounded source} \stm{give a more specific definition}. \ashok{The linear memory is important for unbound astrophysical systems and have been studied in context of supernova explosions resulting in neutron star kicks \cite{Burrows1996}\cite{Muller2004}\stm{put one or more citations after each item}, asymmetric mass loss due to neutrino emission \cite{COtt2009} \stm{isn't this also in the context of supernovae? if not, what source is this?}, and gamma-ray-burst jets \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}}.

@@

This wave equation has several source terms. The Landau-Lifshitz term represents the contribution of GWs to the stress tensor,
and therefore serves as the source of the nonlinear memory contribution to GW waveform \cite{Isaacson1968},
\stm{You appear to change the meanings of T and t, which you shouldn't do, but please verify that my correction is what you actually meant. Also, since the contravariant form appears in the above equation, you should give that here, rather than the covariant form.}
\begin{equation}
	t_{LL}^{jk}\equiv \frac{1}{32\pi}\langle h_{TT}^{ab,j}h_{TT}^{ab,k}\rangle\approx t_{LL}^{00}n^{j}n^{k} = \frac{1}{R^{2}}\frac{dE^{gw}}{dt d\Omega}n^{j}n^{k}
\end{equation}
where $TT$ refers to the transverse-traceless gauge, $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-brackets correspond to an average over several wavelengths, \stm{what does this last one mean?} $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$. The standard procedure to solve a wave equation with a source is to use the Green's function. When this approach is applied to Eq. \ref{eq:3.1a}, we obtain a correction term to the gravitational field sourced by the Landau-Lifshitz term, given by \cite{Wiseman1991}
\begin{equation}\label{eq:1.6}
	\delta h^{TT}_{jk} = \frac{4}{R}\int_{-\infty}^{T_{R}} dt'\Bigg[ \frac{dE^{gw}}{dt'd\Omega'}\frac{n'_{j}n'_{k}}{(1-\boldsymbol{n'}\cdot\boldsymbol{N})}d\Omega'\Bigg]^{TT}
\end{equation}
where $T_{R}$ is the retarded time. Eq.~\ref{eq:1.6} is the nonlinear Landau-Lifshitz pseudotensor describing gravitational wave memory \cite{Christodoulou1991, Payne1894, Blanchet_Damour1992}. The nonlinear memory occurs in any system that radiates GWs, whereas these systems are usually considered to have a vanishing linear memory, although we can apply the formalism for linear memory to nonlinear memory if 
we consider the radiated gravitons to be the unbound masses \cite{Thorne1992}.
%The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992}

@@

\bibliographystyle{apsrev4-1}
\bibliography{references}
